\textls[-0]{\thetitle}

### 2023-03-06 canonicitymodel.tex
Working on Canonicity Model.

Like most logical relation proof,  we interpret each judgement, inductively interpret each syntax piece and prove compatibility lemma. We are working on an intrinsic setting thus even we omit context for brevity, the syntax piece is still well-typed. We only specify part of the logical relation proof for MLTT part. For the complete version, please refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic}.

\begin{align*}
  \denotesC{\goodCtx{\Gm}{i}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
  % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
  \denotesC{\goodType{\Gm}{T}{}} & \text{ is a function maps to Set s.t.} \\
  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a Set of closed terms } \goodTerm{\cdot}{\_}{\sub{T}{\gm}} \\ 
  % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
  \denotesC{\goodSub{\Gm}{\gm}{\Delta}} & \text{ is a function}: \denotesC{\goodCtx{\Gm}{i}} \to \denotesC{\goodCtx{\Delta}{i}} \text{ s.t. } \denotesC{\goodSub{\Gm}{\gm}{\Delta}}(\sigma) \coloneqq \SubstComp{\sigma}{\gm} \\
  % &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
  \denotesC{\goodTerm{\Gm}{t}{T}} & \text{ is a dependent function}: \prod_{\gm \in \denotesC{\goodCtx{\Gm}{i}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)  \text{ s.t. } \denotesC{\goodTerm{\Gm}{t}{T}}(\gm) \coloneqq \sub{t}{\gm} \\
  \denotesC{\goodType{\Gm}{\top}{}}(\gm) &= \{t : t \equiv \unit \} \\
  \denotesC{\goodType{\Gm}{\bot}{}}(\gm) &= \emptyset \\
  \denotesC{\goodType{\Gm}{\cB}{}}(\gm) &= \{t : t \equiv \true \text{ or } t \equiv\false \} \\
  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm) &= \{t : t \equiv \eqrefl{\sub{a}{\gm}} \} \text{ when } \sub{a}{\gm} \equiv \sub{b}{\gm} \\ 
  &= \emptyset \text{ otherwise } \\ 
  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}} &= \\
  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}} &= \\
  \denotesC{\goodCtx{\cdot}{}} &= \{\SubstEmp\} \text{  the singleton set with identity substitution} \\ 
  \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
  \denotesC{\goodTerm{\Gm}{\unit}{\top}} &= \\
  \denotesC{\goodTerm{\Gm}{\true}{\cB}} &= \\
  \denotesC{\goodTerm{\Gm}{\false}{\cB}} &= \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}} &= \\
  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}} &= \\
  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\app{t}}{B}} &= \\
  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}} &= \\
  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}} &= \\
\end{align*}

@@

Once we have above model for MLTT, we can notice that we have a function $\Glued{\Pi}$ s.t. if $A' = \denotesC{A}, B' = \denotesC{B}$ then $\Glued{\Pi}(A', B') = \denotesC{\TyPi{A}{B}}$. Using more type-theoretical language, this $\Glued{\Pi}$ is using the \emph{internal dependent function type} of the above model. Similarly, we can have $\Glued{\Sigma}$ and $\GSubstExt{\gm}{t}$ for dependent tuple type and substitution extension. These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.

Then we extend the above model to \TT , we still start with the judgement. Since linkage transformers are just syntactic sugar, we will not interpret it.
The intuition of the model is similar to that of the syntactic translation---use the canonicity model of sigma type (from the earlier work) for linkage type and using inductive facility in the meta-logic to justify our W-type.

\begin{align*}
  \denotesC{\goodSig{\Gm}{\lsig}{n}} & \text{ is a list of 3-tuple of length $n$  } \\
  % & \text{ and we define  } {\goodSigC{\Gm}{\lsig}{n}} :\iff \denotesC{\goodSig{\Gm}{\lsig}{n}} \text{ is defined} \\ 
  \denotesC{\goodWSig{\Gm}{\wsig}{n}} &\text{ is a list of 2-tuple of length $n$ } \\
  % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
  & \text{we define }\TyLkg{}, \TyTkg{} \text{ inductively on its input signature} \\ 
  \mathit{\cL}("nil") &= \denotesC{\top} \\
  \mathit{\cL}((A,s,T) :: tl)(\gm) &= {\mathit{\cL}(tl)}(\gm) \times \mathit{\Pi}(A,T)(\gm) \\
  \mathit{\cP}("nil") &= \denotesC{\top}\\
  \mathit{\cP}((A,s,T) :: tl) &= \mathit{\Sigma}(\mathit{\cP}(tl), \sub{T}{(p_1, s)})\\
  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \mathit{\cL}(\denotesC{\lsig}) \\
  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \mathit{\cP}(\denotesC{\lsig}) \\
  % 
  \denotesC{\goodSig{\Gm}{\LSigEmp}{0}} &= "nil"\\ 
  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesC{A}, \denotesC{s}, \denotesC{T})::\denotesC{\lsig}\\ 
  \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[0]\\ 
  \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= ("hd"~\denotesC{\lsig})[1]\\ 
  \denotesC{\goodType{\Gm, \lsigprojT{\lsig}}{\lsigproj{2}{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[2]\\
  \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &=  \\
  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}} &= \\
  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\ell}}{\TyTkg{\lsig}}} &= \\
  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\ell}}{T}} &= \\
  \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= \\
  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= \\
  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= \\
  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= \\
  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{}} &= \\
  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
  \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= \\
  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= {\mathit{\Pi}{A}{\mathit{\Pi} {\mathit{\Pi}{B}{\sub{R}{\SubstWeak{2}}}} {\sub{R}{\SubstWeak{2}}}}}\\
  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= \\
  \denotesC{\Recproj{j}{\lkg}} &= \\
  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}} &= \\
\end{align*}

>>>>

Like most logical relation proof,  we interpret each judgement, inductively interpret each syntax piece. We are working on an intrinsic setting thus even we omit context for brevity, the syntax piece is still well-typed. We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model). Interested audience can refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} for a complete technical treatements.

\begin{align*}
  \denotesC{\goodCtx{\Gm}{i}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
  % &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
  \denotesC{\goodType{\Gm}{T}{}} & \text{ is a function maps to Set s.t.} \\
  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a Set of closed terms } \goodTerm{\cdot}{\_}{\sub{T}{\gm}} \\ 
  % &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
  \denotesC{\goodSub{\Gm}{\gm}{\Delta}} & \text{ is a function}: \denotesC{\goodCtx{\Gm}{i}} \to \denotesC{\goodCtx{\Delta}{i}} \text{ s.t. } \denotesC{\goodSub{\Gm}{\gm}{\Delta}}(\sigma) \coloneqq \SubstComp{\sigma}{\gm} \\
  % &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
  \denotesC{\goodTerm{\Gm}{t}{T}} & \text{ is a dependent function}: \prod_{\gm \in \denotesC{\goodCtx{\Gm}{i}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)  \text{ s.t. } \denotesC{\goodTerm{\Gm}{t}{T}}(\gm) \coloneqq \sub{t}{\gm} \\
  \denotesC{\goodType{\Gm}{\top}{}}(\gm) &= \{t : t \equiv \unit \} \\
  \denotesC{\goodType{\Gm}{\bot}{}}(\gm) &= \emptyset \\
  \denotesC{\goodType{\Gm}{\cB}{}}(\gm) &= \{t : t \equiv \true \text{ or } t \equiv\false \} \\
  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm) &= \{t : t \equiv \eqrefl{\sub{a}{\gm}} \} \text{ when } \sub{a}{\gm} \equiv \sub{b}{\gm} \\ 
  &= \emptyset \text{ otherwise } \\ 
  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}} &= \\
  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}} &= \\
  \denotesC{\goodCtx{\cdot}{}} &= \{\SubstEmp\} \text{  the singleton set with identity substitution} \\ 
  \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
  \denotesC{\goodTerm{\Gm}{\unit}{\top}}(\gm) &= \unit\\
  \denotesC{\goodTerm{\Gm}{\true}{\cB}}(\gm) &= \true \\
  \denotesC{\goodTerm{\Gm}{\false}{\cB}}(\gm) &= \false \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}}(\gm) &= \\
  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}}(\gm) &= \eqrefl{\sub{t}{\gm}}\\
  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\app{t}}{B}} &= \\
  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}} &= \\
  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}} &= \\
\end{align*}

@@

Once we have above model for MLTT, we can notice that we have a function $\Glued{\Pi}$ s.t. if $A' = \denotesC{A}, B' = \denotesC{B}$ then $\Glued{\Pi}(A', B') = \denotesC{\TyPi{A}{B}}$. Type-theoretical speaking, this $\Glued{\Pi}$ is using the \emph{internal dependent function type} of the above model. Similarly, we can have $\Glued{\Sigma}$ and $\GSubstExt{\gm}{t}$ for dependent tuple type and substitution extension. These will be useful in our \TT model because as syntactic translation hinted, \TT reuses a lot from MLTT.

Now we can extend the above model to \TT , we still start with the judgement. Since linkage transformers are just syntactic sugar, we will not interpret it.
The intuition of the model is similar to that of the syntactic translation---use the canonicity model of sigma type (from the earlier work) for linkage type and using inductive facility in the meta-logic to justify our W-type.

\begin{align*}
  \denotesC{\goodSig{\Gm}{\lsig}{n}} & \text{ is a list of 3-tuple of length $n$  } \\
  % & \text{ and we define  } {\goodSigC{\Gm}{\lsig}{n}} :\iff \denotesC{\goodSig{\Gm}{\lsig}{n}} \text{ is defined} \\ 
  \denotesC{\goodWSig{\Gm}{\wsig}{n}} &\text{ is a list of 2-tuple of length $n$ } \\
  % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
  & \text{we define }\TyLkg{}, \TyTkg{} \text{ inductively on its input signature} \\ 
  \mathit{\cL}("nil") &= \denotesC{\top} \\
  \mathit{\cL}((A,s,T) :: tl)(\gm) &= {\mathit{\cL}(tl)}(\gm) \times \mathit{\Pi}(A,T)(\gm) \\
  \mathit{\cP}("nil") &= \denotesC{\top}\\
  \mathit{\cP}((A,s,T) :: tl) &= \mathit{\Sigma}(\mathit{\cP}(tl), \sub{T}{(p_1, s)})\\
  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= \mathit{\cL}(\denotesC{\lsig}) \\
  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \mathit{\cP}(\denotesC{\lsig}) \\
  % 
  \denotesC{\goodSig{\Gm}{\LSigEmp}{0}} &= "nil"\\ 
  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesC{A}, \denotesC{s}, \denotesC{T})::\denotesC{\lsig}\\ 
  \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[0]\\ 
  \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= ("hd"~\denotesC{\lsig})[1]\\ 
  \denotesC{\goodType{\Gm, \lsigprojT{\lsig}}{\lsigproj{2}{\lsig}}{}} &= ("hd"~\denotesC{\lsig})[2]\\
  \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm) &= (\denotesC{\ell}(\gm), \lambda u . t (\gm, u)) \\
  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \\
  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\ell}}{\TyTkg{\lsig}}} &= \\
  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\ell}}{T}} &= \\
  \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\
  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= (\denotesC{A} , \denotesC{B}) :: \denotesC{\wsig}\\
  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[0]\\
  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= (\text{ $j$-th element of }~\denotesC{\wsig})[1]\\
  \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= "tl"~\denotesC{\wsig}\\
  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}} &= \\
  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}} &= \\
  \denotesC{\goodType{\Gm}{\CaseSig{A}{B}{R}}{}} &= {\mathit{\Pi}{A}{\mathit{\Pi} {\mathit{\Pi}{B}{\sub{R}{\SubstWeak{2}}}} {\sub{R}{\SubstWeak{2}}}}}\\
  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= \\
  \denotesC{\Recproj{j}{\lkg}} &= \\
  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}} &= \\
\end{align*}

### 2023-03-06 metatheory2.tex
Working on Canonicity Model.

\subsection{Standard Model for Consistency}

\input{standardmodel.tex}

@@

So we can inductively define the syntax sugar $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ using constructors $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$. Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality given in the syntax. In that way, the explicit substituon for linkage transformer becomes meta-level substitution again; and $\inh{h}{\ell}$ is not itself a syntax but a meta-level computation that leads to a syntax.

>>>>

% \subsection{Standard Model for Consistency}

% \input{standardmodel.tex}

@@

So we can inductively define the syntax sugar $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$ using constructors $\InhExt{}{}$, $\InhOv{}{}$, $\InhInh{}$, $\InhNest{}{}$. Then we define the $\inh{h}{\ell}$ and the explicit substitution for linkage transformer $\sub{h}{\gm}$ by induction on the structure of linkage transformer $h$, based on the equality\EDJ{Add substitution equalities for linkage transformers} given in the syntax. In that way, the explicit substituon for linkage transformer becomes meta-level substitution again; and $\inh{h}{\ell}$ is not itself a syntax but a meta-level computation that leads to a syntax.
